Identifying and correcting for misspecifications\\in GWAS summary statistics for polygenic scores

### 2022-02-18 paper/paper-misspec-supp.tex
add mismatched LD
save

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{compare-mult-imp}}
	\caption{Corrected GWAS \textbf{A:} Z-scores and \textbf{B:} effect sizes from imputed dosages (multiplying them by $\text{INFO}$) versus the ones from multiple imputation (MI, 10 draws used), colored by INFO scores. This shows only one tenth of the simulated variants, because MI can be computationally intensive. \label{fig:compare-mi}}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=0.95\textwidth]{simu-rounded}}
	\caption{Results for the simulations with summary statistics (effect sizes and their standard errors) possibly rounded to two significant digits, averaged over 10 simulations for each scenario. Reported 95\% confidence intervals are computed from 10,000 non-parametric bootstrap replicates of the mean. Red bars correspond to using the LD with independent blocks (Methods).}
	\label{fig:simu-rounded}
\end{figure}

>>>>

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{compare-mult-imp}}
	\caption{(Over)corrected GWAS \textbf{A:} Z-scores and \textbf{B:} effect sizes from imputed dosages (multiplying them by $\text{INFO}$) versus the ones obtained from multiple imputation (MI, 20 draws used, \cite{Palmer2016}), colored by INFO scores. This shows only one tenth of the simulated variants, because MI can be computationally intensive. \label{fig:compare-mi}}
\end{figure}

@@

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{simu-rounded}}
	\caption{Results for the simulations with summary statistics (effect sizes and their standard errors) possibly rounded to two significant digits, averaged over 10 simulations for each scenario. Reported 95\% confidence intervals are computed from 10,000 non-parametric bootstrap replicates of the mean. Red bars correspond to using the LD with independent blocks (Methods).}
	\label{fig:simu-rounded}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{simu-altpop}}
	\caption{Results for the simulations with summary statistics with LD matrices based on two different populations. One comes from the same population used for computing the GWAS summary statistics (North-West Europe), while the other one comes from South Europe (alternative LD reference). Reported 95\% confidence intervals are computed from 10,000 non-parametric bootstrap replicates of the mean. Red bars correspond to using the LD with independent blocks (Methods).}
	\label{fig:simu-altpop}
\end{figure}

### 2022-02-18 paper/paper-misspec.tex
save

Contrary to individual-level genotypes and phenotypes, summary statistics resulting from genome-wide association studies (GWAS) are widely available, and very large sample sizes can be obtained through meta-analysis \cite[]{yengo2022saturated}.
Therefore, they have been extensively used to e.g.\ derive polygenic scores (PGS), perform fine-mapping, and estimate key genetic architecture parameters.
However, GWAS summary statistics come with uneven quality, such as the imputation accuracy of each variant reported. 
There is also heterogeneity in the information made available in these summary statistics, e.g.\ per-variant imputation INFO scores, sample sizes and allele frequencies are often missing.
Moreover, many methods based on summary statistics use Bayesian models, which can be sensitive to model misspecifications \cite[]{WALKER20131621,miller2018robust}.

We present an overview of possible misspecifications that come with GWAS summary statistics (Table \ref{tab:overview}). 
First, the total sample size used can be misestimated, which would result e.g.\ in a biased estimation of the SNP heritability. For example, the total effective sample size is overestimated when computed from the total number of cases and controls from a meta-analysis of binary outcomes \cite[]{grotzinger2021pervasive}; using BOLT-LMM summary statistics can result in an increased effective sample size \cite[]{loh2018mixed,gazal2018functional}; using SAIGE on binary traits with a large prevalence can result in a reduced effective sample size \cite[]{zhou2018efficiently,mbatchou2021computationally}. 
Second, per-variant sample sizes can be very different when meta-analyzing GWAS summary statistics from multiple cohorts with different sets of variants \cite[]{wang2021global}.
Per-variant sample sizes should be used instead of the overall total sample size to avoid any bias resulting from overestimating the sample size.
Moreover, using very different per-variant sample sizes could be problematic for models implicitly assuming that summary statistics have all been derived from the same individuals \cite[]{zhu2017bayesian,zhou2021fast}.
As a quality control (QC) step, one can e.g.\ remove SNPs with a sample size less than 0.67 times the 90th percentile of all sample sizes \cite[]{zheng2017ld}.
Third, many genetic variants are imputed, which can cause some bias in summary statistics.
Fourth, the quality of imputation of each variant, often used in a QC step in statistical genetics analyses, can be severely misestimated (often overestimated) when computed for multi-ancestry individuals instead of a more homogeneous subset.
Fifth, errors such as allele inversions can be present in the summary statistics; QC is particularly important here.
Sixth, there can be some mismatch between the GWAS summary statistics and the LD reference used.

@@

Here we investigate some of these misspecifications, and propose adjustments to improve the predictive performance of polygenic scores derived from GWAS summary statistics.
We approach this from three different angles.
First, based on additional summary information such as the imputation INFO scores and allele frequencies from the GWAS summary statistics, we refine our previously proposed QC \cite[]{prive2020ldpred2}.
This QC consists in comparing standard deviations (of genotypes) inferred from GWAS summary statistics with the ones computed from a reference panel.
This is useful to check that the input parameters used are consistent with each other. 
This was particularly important for LDpred2-auto, which directly estimates key model parameters from the data.
Here, we show e.g.\ that standard deviations of imputed genotypes (allele dosages) are lower than the expected values under Hardy--Weinberg equilibrium.
Second, we investigate possible adjustments to apply to the input parameters of methods using summary statistics (focusing on PGS methods here), namely the reference LD (linkage disequilibrium) matrix, the GWAS effect sizes, their standard errors and the corresponding sample sizes.
For example, we show that GWAS effect sizes computed from imputed dosages are larger in magnitude compared to if true genotypes were available.
Third, we introduce two new optional parameters in LDpred2-auto to make it more robust to these types of misspecification.
We focus our investigations on LDpred2 and lassosum for two reasons. First, multiple studies have shown that LDpred2 and lassosum rank among the best methods for single-trait polygenic prediction \cite[]{mak2017polygenic,prive2020ldpred2,pain2021evaluation,kulm2021systematic}.
Second, we reimplement and use a new version of lassosum, called lassosum2, that uses the exact same input parameters as LDpred2, which makes it easy for us to test the different QCs and adjustments presented here.

@@

We design simulations where variants have different GWAS sample sizes, which is often the case when meta-analyzing GWAS summary statistics from multiple cohorts with different sets of variants \cite[]{wang2021global}.
Using 40,000 variants from chromosome 22 (Methods), we simulate quantitative phenotypes with a heritability of 20\% and 2000 causal variants.
We then divide the 40,000 variants into three groups: for half of the variants, we use 100\% of 300,000 individuals for GWAS, but use only 80\% for one quarter of the variants and 60\% for the remaining quarter. 
We then run C+T, lassosum, lassosum2 (Methods), LDpred2-inf, LDpred2(-grid), and LDpred2-auto \cite[]{prive2019making,mak2017polygenic,prive2020ldpred2} by using either the true per-variant GWAS sample sizes, the total sample size, or imputed sample sizes. 
Note that we initially included PRS-CS and SBayesR in our comparison \cite[]{ge2019polygenic,lloyd2019improved}. However, results for SBayesR always diverged (stopped with an error), and the overlap with the LD reference provided for PRS-CS was too small.
Averaged over 10 simulations, when providing true per-variant GWAS sample sizes, squared correlations between the polygenic scores and the simulated phenotypes are of 0.123 for C+T, 0.161 for lassosum, 0.169 for lassosum2, 0.159 for LDpred2(-grid), 0.140 for LDpred2-auto, and 0.141 for LDpred2-inf (Figure \ref{fig:simu-misN}).
Results when using imputed sample sizes are quite similar to when using the true ones.
Note that C+T does not use this sample size information.
When using the total GWAS sample size instead of the per-variant sample sizes, predictive performance slightly decreases to 0.157 for lassosum and to 0.163 for lassosum2, but dramatically decreases for LDpred2 with new values of 0.134 for LDpred2-grid, 0.119 for LDpred2-auto, and 0.123 for LDpred2-inf (Figure \ref{fig:simu-misN}).
This extreme simulation scenario shows that LDpred2 can be sensitive to GWAS sample size misspecification, whereas lassosum (and lassosum2) seems little affected by this.

@@

Here we design simulations to understand the impact of using a mismatched LD reference panel, e.g.\ that comes from a different population compared to the one used to compute the GWAS summary statistics.
We use the same simulation setup as before (Methods).
In addition, we design an alternative reference panel based on 10,000 individuals from South Europe by using the ``Italy'' center defined in \cite{prive2021high}.
Averaged over 10 simulations, when using an ``alternative'' LD reference panel instead of a well-matched one, squared correlations between the polygenic scores and the simulated phenotypes drop from 0.174 to 0.167 for lassosum2, from 0.168 to 0.147 for LDpred2(-grid), from 0.174 to 0.169 for LDpred2-low-h2, from 0.143 to 0.139 for LDpred2-auto, from 0.162 to 0.146 for LDpred2-auto-rob, and from 0.144 to 0.140 for LDpred2-inf (Figure \ref{fig:simu-altpop}).
Defining LD blocks in these two LD matrices always helps, but do not improve performance by much when using the alternative LD.

@@

Breast cancer summary statistics are interesting because they include results from two mega analyses \cite[]{michailidou2013large,michailidou2015genome,michailidou2017association}, which means that parameters reported in the GWAS summary statistics, such as the INFO scores and the sample sizes, are estimated with high precision.
Imputation INFO scores for the OncoArray summary statistics (after having restricted to HapMap3 variants) are generally very good (Figure \ref{fig:hist-info-brca}) and better than the ones from iCOGS (Figure \ref{fig:hist-info-brca2}), probably because the iCOGS chip included around 200K variants only, compared to more than 500K variants for the OncoArray.
For both summary statistics, we first compare the standard deviations (of genotypes) inferred from the reported allele frequencies (i.e.\ $\sqrt{2 f (1 - f)}$ where $f$ is the allele frequency, and denoted as sd\_af) versus the ones inferred from the GWAS summary statistics (Equation \eqref{eq:approx-sd-log}, and denoted as sd\_ss).
When coloring by INFO scores, we see a clear trend with sd\_ss being lower than sd\_af as INFO decreases; indeed, using $\text{sd\_ss} / \sqrt{\text{INFO}}$ provides a very good fit for sd\_af, except for some variants of chromosome 6 and 8 for the OncoArray summary statistics (Figures \ref{fig:brca_onco_qc} and \ref{fig:brca_icogs_qc}). Most of these outlier variants are in regions 25-33 Mbp of chromosome 6 and 8-12 Mbp of chromosome 8 (Figure \ref{fig:hist-bad-pos}), which are two known long-range LD regions \cite[]{price2008long}.
We hypothesize that this is due to using (as covariates in the GWAS) principal components (PCs)  that capture LD structure instead of population structure \cite[]{prive2020efficient}.
To validate this hypothesis, we simulate a phenotype using HapMap3 variants of chromosome 6 for 10,000 individuals from the UK Biobank, then we run GWAS with or without PC19 as covariate. PC19 from the UK Biobank was previously reported to capture LD structure in region 70-91 Mbp of chromsome 6 \cite[]{prive2020efficient}.
In these simulations, the same bias as in figure \ref{fig:brca_onco_qc}B is observed for the variants in this region (Figure \ref{fig:gwas_bad_pc}), confirming our hypothesis.

@@

Therefore, providing an accurate imputation INFO score is useful for two reasons. First, it allows for correcting for a reduced standard deviation when using imputed data in the QC step we propose, in order to better uncover problems with the summary statistics.
Second, using one of the proposed corrections based on the INFO scores may lead to an improved prediction when deriving polygenic scores.
We apply this correction to the two breast cancer summary statistics. We first compare the standard QC proposed in \cite{prive2020ldpred2} (which ends up filtering on MAF here, which we call ``qc1''). We then also filter out the two long-range LD regions of chromosome 6 and 8 for the OncoArray summary statistics and remove around 500 variants when filtering on differences of MAFs between summary statistics and the validation dataset (``qc2'').
As for the correction for the INFO scores, we use the first correction, ``sqrt\_info'', which is simple because it is equivalent to post-processing PGS effects by multiplying them by $\sqrt{\text{INFO}}$.
Although results for both QC used are very similar, correcting for the INFO score slightly improves prediction when deriving polygenic scores based on iCOGS summary statistics (Figure \ref{fig:res_brca}). 
All other improvements introduced before have little to no effect here, probably because misspecifications are much smaller than in the simulations.

@@

Note that, in this study, we use summary statistics that include extended information (e.g.\ INFO scores and allele frequencies), yet most GWAS summary statistics do not provide such exhaustive information \cite[]{macarthur2021workshop}.
We acknowledge that, in the case of a meta-analysis from multiple studies, providing a single INFO score per variant may not be possible; would a weighted averaged INFO score work?
Nevertheless, this quality control could be performed within each study before pooling results, to make sure that summary statistics have the best possible quality for follow-up analyses such as deriving polygenic scores.
Another information, the effective sample size per variant, is often missing from GWAS summary statistics. 
Sometimes, it can even be challenging to recover the total effective sample size from large meta-analyses.
We recall that if some studies of a meta-analysis have an imbalanced number of cases and controls, the total effective sample size should not be computed from the total numbers of cases and controls overall, but instead from the sum of the effective sample sizes of each study.
Indeed, take the extreme example of meta-analyzing two studies, one with 1000 cases and 0 controls, and another one with 0 cases and 1000 controls, then the effective sample size of the meta-analysis is 0, not 2000.
This misspecification can lead to serious issues such as misestimating the SNP heritability \cite[]{grotzinger2021pervasive}.
Fortunately, an overestimated sample size can be detected from the QC plot we propose, where the slope is then less than 1 for case-control studies using logistic regression; otherwise the standard deviation of the phenotype is also needed (Equation \ref{eq:approx-sd-lin}), but can be estimated \cite[]{prive2020ldpred2}.

@@

Nevertheless, we recommend to adopt these changes, i.e.\ performing the (refined) QC proposed here, defining LD blocks in the LD matrix, and using more regularization when needed (i.e.\ also testing smaller values for the heritability parameter in LDpred2-grid, and using the two new parameters in LDpred2-auto (Section \ref{sec:robust}). 
Note that LD blocks are already widely used by several methods (e.g.\ lassosum and PRS-CS) because they allow for processing smaller matrices at once \cite[]{mak2017polygenic,ge2019polygenic}; here we have shown that these blocks are also useful to make methods more robust.
PRS-CS is currently one of the most robust PGS methods; for example, it can use the (small) 1000 Genomes data as LD reference \cite[]{ge2019polygenic}, and can even use a European LD reference panel with multi-ancestry GWAS summary statistics \cite[]{wang2021global}.
We think this is made possible by the use of a (very) strong regularization in PRS-CS ($\phi^{-1} \psi_j^{-1} \ge 1$, which would approximately correspond to using $s=0.5$ in lassosum, $\delta = 1$ in lassosum2, and $\texttt{shrink\_corr}=0.5$ in LDpred2-auto).
Using enough regularization is good for robustness, but note that using too much regularization can also damage predictive performance, and should not be a replacement for a proper QC and choice of ancestry-matched LD [TODO: A BIT AGGRESSIVE?].
As a future research direction, we are interested in using ancestry-matched LD matrices to improve polygenic prediction in all ancestries.
Also it would be useful to investigate the impact of the QC and adjustments we propose here on other (non-PGS) methods, such as the ones for inference, where robustness is likely to be very important as well.

>>>>

Contrary to individual-level genotypes and phenotypes, summary statistics resulting from genome-wide association studies (GWAS) are widely available, and very large sample sizes can be obtained through meta-analysis \cite[]{yengo2022saturated}.
GWAS summary statistics have been extensively used to derive polygenic scores (PGS), perform fine-mapping, and estimate a range of key genetic architecture parameters.
However, GWAS summary statistics come with uneven imputation accuracy.
There is also heterogeneity in the information made available in these summary statistics; for example, per-variant imputation INFO scores, sample sizes and allele frequencies are often missing.
Moreover, many methods based on summary statistics use Bayesian models, which can be sensitive to model misspecifications \cite[]{WALKER20131621,miller2018robust}.

We present an overview of possible misspecifications that come with GWAS summary statistics (Table \ref{tab:overview}). 
First, the total sample size used can be misestimated, which would result in a biased estimation of the SNP heritability. For example, the total effective sample size is overestimated when computed from the total number of cases and controls from a meta-analysis of binary outcomes \cite[]{grotzinger2021pervasive}; using BOLT-LMM summary statistics can result in an increased effective sample size \cite[]{loh2018mixed,gazal2018functional}; using SAIGE on binary traits with a large prevalence can result in a reduced effective sample size \cite[]{zhou2018efficiently,mbatchou2021computationally}. 
Second, per-variant sample sizes can vary substantially and be much smaller than the total sample size when meta-analyzing GWAS summary statistics from multiple cohorts with different sets of variants \cite[]{wang2021global}.
Using very different per-variant sample sizes can be problematic for models implicitly assuming that summary statistics have all been derived from the same individuals \cite[]{zhu2017bayesian,zhou2021fast}.
As a quality control (QC) step, one can e.g.\ remove SNPs with a sample size less than 0.67 times the 90th percentile of all sample sizes \cite[]{zheng2017ld}.
Third, many genetic variants are imputed, which can cause some bias in summary statistics.
Fourth, the imputation quality of each variant (e.g.\ INFO scores), often used in a QC step in statistical genetics analyses, can be severely misestimated (often overestimated) when computed for multi-ancestry individuals instead of a more homogeneous subset.
Fifth, errors such as allele inversions can be present in the summary statistics; QC is particularly important here.
Sixth, there can be some mismatch between the GWAS summary statistics and the LD reference used.

@@

Here we investigate some of these misspecifications, and propose adjustments to improve the predictive performance of polygenic scores derived from GWAS summary statistics.
We approach this from three different angles.
First, based on additional summary information such as the imputation INFO scores and allele frequencies from the GWAS summary statistics, we refine our previously proposed QC \cite[]{prive2020ldpred2}.
This QC consists in comparing standard deviations (of genotypes) inferred from GWAS summary statistics with the ones computed from a reference panel.
This is useful to check that the input parameters used are consistent with each other. 
This was particularly important for LDpred2-auto, which directly estimates key model parameters from the data.
Here we further show that standard deviations of imputed genotypes (allele dosages) are lower than the expected values under Hardy--Weinberg equilibrium.
Second, we investigate possible adjustments to apply to the input parameters of methods using summary statistics (focusing on PGS methods here), namely the reference LD (linkage disequilibrium) matrix, the GWAS effect sizes, their standard errors and the corresponding sample sizes.
For example, we show that GWAS effect sizes computed from imputed dosages are larger in magnitude compared to if true genotypes were available.
Third, we introduce two new optional parameters in LDpred2-auto to make it more robust to these types of misspecification.
We focus our investigations on LDpred2 and lassosum for two reasons. First, multiple studies have shown that LDpred2 and lassosum rank among the best methods for single-trait polygenic prediction \cite[]{mak2017polygenic,prive2020ldpred2,pain2021evaluation,kulm2021systematic}.
Second, we reimplement and use a new version of lassosum, called lassosum2, that uses the exact same input parameters as LDpred2, which makes it easy for us to test the different QCs and adjustments presented here.

@@

We design GWAS simulations where variants have different sample sizes, which is often the case when meta-analyzing GWAS summary statistics from multiple cohorts with different sets of variants \cite[]{wang2021global}.
Using 40,000 variants from chromosome 22 (Methods), we simulate quantitative phenotypes with a heritability of 20\% and 2000 causal variants.
We then divide the 40,000 variants into three groups: for half of the variants, we use 100\% of 300,000 individuals for GWAS, but use only 80\% for one quarter of the variants and 60\% for the remaining quarter. 
We then run C+T, lassosum, lassosum2 (Methods), LDpred2-inf, LDpred2(-grid), and LDpred2-auto \cite[]{prive2019making,mak2017polygenic,prive2020ldpred2} by using either the true per-variant GWAS sample sizes, the total sample size, or imputed sample sizes. 
Note that we initially included PRS-CS and SBayesR in our comparison \cite[]{ge2019polygenic,lloyd2019improved}. However, results for SBayesR always diverged (stopped with an error), and the overlap with the LD reference provided for PRS-CS was too small.
When providing true per-variant GWAS sample sizes, squared correlations between the polygenic scores and the simulated phenotypes are of 0.123 for C+T, 0.161 for lassosum, 0.169 for lassosum2, 0.159 for LDpred2(-grid), 0.140 for LDpred2-auto, and 0.141 for LDpred2-inf (Figure \ref{fig:simu-misN}, averaging over 10 simulations).
Results when using imputed sample sizes are quite similar to when using the true ones.
Note that C+T does not use this sample size information.
When using the total GWAS sample size instead of the per-variant sample sizes, predictive performance slightly decreases to 0.157 for lassosum and to 0.163 for lassosum2, but dramatically decreases for LDpred2 with new values of 0.134 for LDpred2-grid, 0.119 for LDpred2-auto, and 0.123 for LDpred2-inf (Figure \ref{fig:simu-misN}).
This extreme simulation scenario shows that LDpred2 can be sensitive to GWAS sample size misspecification, whereas lassosum (and lassosum2) seems little affected by this.

@@

Here we design simulations to understand the impact of using a mismatched LD reference panel, e.g.\ that comes from a different population compared to the one used to compute the GWAS summary statistics.
We use the same simulation setup as before (Methods).
In addition, we design an alternative reference panel based on 10,000 individuals from South Europe by using the ``Italy'' center defined in \cite{prive2021high}.
When using an ``alternative'' LD reference panel instead of a well-matched one, squared correlations between the polygenic scores and the simulated phenotypes drop from 0.174 to 0.167 for lassosum2, from 0.168 to 0.147 for LDpred2(-grid), from 0.174 to 0.169 for LDpred2-low-h2, from 0.143 to 0.139 for LDpred2-auto, from 0.162 to 0.146 for LDpred2-auto-rob, and from 0.144 to 0.140 for LDpred2-inf (Figure \ref{fig:simu-altpop}, averaging over 10 simulations).
Defining LD blocks in these two LD matrices always helps, but do not improve performance by much when using the alternative LD.

@@

Breast cancer summary statistics are interesting because they include results from two mega analyses \cite[]{michailidou2013large,michailidou2015genome,michailidou2017association}, which means that parameters reported in the GWAS summary statistics, such as the INFO scores and the sample sizes, are estimated with high precision.
Imputation INFO scores for the OncoArray summary statistics (after having restricted to HapMap3 variants) are generally very good (Figure \ref{fig:hist-info-brca}) and better than the ones from iCOGS (Figure \ref{fig:hist-info-brca2}), probably because the iCOGS chip included around 200K variants only, compared to more than 500K variants for the OncoArray.
For both summary statistics, we compare the standard deviations (of genotypes) inferred from the reported allele frequencies (i.e.\ $\sqrt{2 f (1 - f)}$ where $f$ is the allele frequency, and denoted as sd\_af) versus the ones inferred from the GWAS summary statistics (Equation \eqref{eq:approx-sd-log}, and denoted as sd\_ss).
As shown in figures \ref{fig:brca_onco_qc} and \ref{fig:brca_icogs_qc}, there is a clear trend with sd\_ss being lower than sd\_af as INFO decreases; indeed, using $\text{sd\_ss} / \sqrt{\text{INFO}}$ provides a very good fit for sd\_af, except for some variants of chromosome 6 and 8 for the OncoArray summary statistics. Most of these outlier variants are in regions 25-33 Mbp of chromosome 6 and 8-12 Mbp of chromosome 8 (Figure \ref{fig:hist-bad-pos}), which are two known long-range LD regions \cite[]{price2008long}.
We hypothesize that this is due to using (as covariates in the GWAS) principal components (PCs)  that capture LD structure instead of population structure \cite[]{prive2020efficient}.
To validate this hypothesis, we simulate a phenotype using HapMap3 variants of chromosome 6 for 10,000 individuals from the UK Biobank, then we run GWAS with or without PC19 as covariate. PC19 from the UK Biobank was previously reported to capture LD structure in region 70-91 Mbp of chromsome 6 \cite[]{prive2020efficient}.
In these simulations, the same bias as in figure \ref{fig:brca_onco_qc}B is observed for the variants in this region (Figure \ref{fig:gwas_bad_pc}), confirming our hypothesis.

@@

Therefore, providing an accurate imputation INFO score is useful for two reasons. First, it allows for correcting for a reduced standard deviation when using imputed data in the QC step we propose, in order to better uncover problems with the summary statistics.
Second, using one of the proposed corrections based on the INFO scores may lead to an improved prediction when deriving polygenic scores.
We apply this correction to the two breast cancer summary statistics. We first compare the standard QC proposed in \cite{prive2020ldpred2} (which ends up filtering on MAF here, which we call ``qc1''). We then also filter out the two long-range LD regions of chromosome 6 and 8 for the OncoArray summary statistics and remove around 500 variants when filtering on differences of MAFs between summary statistics and the validation dataset (``qc2'').
As for the correction using the INFO scores, we use the first correction, ``sqrt\_info'', which is simple because it is equivalent to post-processing PGS effects by multiplying them by $\sqrt{\text{INFO}}$.
Although results for both QC used are very similar, correcting using the INFO scores slightly improves prediction when deriving polygenic scores based on iCOGS summary statistics (Figure \ref{fig:res_brca}). 
All other improvements introduced before have little to no effect here, probably because misspecifications are much smaller than in the simulations.

@@

Note that, in this study, we use summary statistics that include extended information (e.g.\ INFO scores and allele frequencies), yet most GWAS summary statistics do not provide such exhaustive information \cite[]{macarthur2021workshop}.
We acknowledge that, in the case of a meta-analysis from multiple studies, providing a single INFO score per variant may not be possible.
Solutions such as using a weighted averaged INFO score might be worth exploring in future studies.
Nevertheless, this quality control could be performed within each study before pooling results, to make sure that summary statistics have the best possible quality for follow-up analyses such as deriving polygenic scores.
Another information, the effective sample size per variant, is often missing from GWAS summary statistics. 
Sometimes, it can even be challenging to recover the total effective sample size from large meta-analyses.
We recall that if some studies of a meta-analysis have an imbalanced number of cases and controls, the total effective sample size should not be computed from the total numbers of cases and controls overall, but instead from the sum of the effective sample sizes of each study.
Indeed, take the extreme example of meta-analyzing two studies, one with 1000 cases and 0 controls, and another one with 0 cases and 1000 controls, then the effective sample size of the meta-analysis is 0, not 2000.
This misspecification can lead to serious issues such as misestimating the SNP heritability \cite[]{grotzinger2021pervasive}.
Fortunately, an overestimated sample size can be detected from the QC plot we propose, where the slope is then less than 1 for case-control studies using logistic regression; otherwise the standard deviation of the phenotype is also needed (Equation \ref{eq:approx-sd-lin}), but can be estimated \cite[]{prive2020ldpred2}.

@@

Nevertheless, we recommend to adopt these changes, i.e.\ performing the (refined) QC proposed here, defining LD blocks in the LD matrix, and using more regularization when needed (i.e.\ also testing smaller values for the heritability parameter in LDpred2-grid, and using the two new parameters in LDpred2-auto (Section \ref{sec:robust}). 
Note that LD blocks are already widely used by several methods (e.g.\ lassosum and PRS-CS) because they allow for processing smaller matrices at once \cite[]{mak2017polygenic,ge2019polygenic}; here we have shown that these blocks are also useful to make methods more robust.
PRS-CS is currently one of the most robust PGS methods; for example, it can use the (small) 1000 Genomes data as LD reference \cite[]{ge2019polygenic}, and can even use a European LD reference panel with multi-ancestry GWAS summary statistics \cite[]{wang2021global}.
We think this is made possible by the use of a (very) strong regularization in PRS-CS ($\phi^{-1} \psi_j^{-1} \ge 1$, which would approximately correspond to using $s=0.5$ in lassosum, $\delta = 1$ in lassosum2, and $\texttt{shrink\_corr}=0.5$ in LDpred2-auto).
Using enough regularization is good for robustness, but note that using too much regularization can also damage predictive performance.
To address this limitation, we recommend to rely on a proper QC and choice of ancestry-matched LD instead of on too much regularization.
As a future research direction, we are interested in using ancestry-matched LD matrices to improve polygenic prediction in all ancestries.
Also it would be useful to investigate the impact of the QC and adjustments we propose here on other (non-PGS) methods, where robustness is likely to be very important as well.
